Forecasting the old age dependency ratio to determine the best pension age

### 2017-02-03 PensionAge.Rmd
work on plane to Perth

---
title: "Forecasting the population age structure to determine the best pension age"
author: "Rob J Hyndman, Yijun Zeng and Han Lin Shang"
date: "`r format(Sys.time(), '%-d %B %Y')`"
fontsize: 11pt
papersize: a4
fontfamily: kpfonts
bibliography: pensionage.bib
biblio-style: authoryear-comp
subparagraph: true
toc: false
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(demography)
library(addb)
library(ggplot2)
source("Rcode/functions-v4.R")
```

@@

In this paper, we use the old aged dependency ratio (OADR\footnote{All abbreviations have been summarized in Appendix A.}) as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working age population to finance pensions and health costs [@ARW07; @IGR02; @IGR07].  The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age [@ARW07]. This can be expressed using the formula below:
\begin{equation}
  OADR=\frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100%
\end{equation}
The research presented here forecasts future population age structures over the next 50 years. We then use those projections to compute the OADR associated with pension age both before and after the adjustments. One major contribution of the current study is the combined use of the Hyndman-Ullah forecasting method [@HU07] and the product-ratio method [@HBY13] to forecast a population age structure.  Forecasts of population structure have been built on forecasts of age and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age and gender-specific population structure.

Using these predicted population paths, we can calculate the average of OADR and predict the levels of old age dependency that correspond to different pension age schemes. By testing different pension age schemes, we aim to find the pension age with the desired OADR outcome. Therefore, other main contributions of this research are (1) to examine the effect of adjusting pension ages by comparing OADR before and after adjustment; and (2)\ to suggest a suitable pension age scheme to ensure that the economic burden will remain relatively constant at a desired level in the future. It should be noted that the desired level has been set at 21%, which is the estimated current level of OADR in 2014. This level was chosen because, while the current financial burden on tax payers is affordable, it is approaching the limits of affordability. However, whether or not the current level is the most appropriate level to Australia remains to be examined and it is not the focus of this paper. This paper concentrates on providing a framework for policy makers to find a target pension age for a given desired level of OADR.

Some may argue that using OADR as a measure of financial burden for the next generation could be biased, with a large number of retirees able to rely on self-financed retirement income, such as superannuation. However, 80% of Australians over pension age fully or partly rely on the age pension [@Power14]. Therefore, the proportion of people over working age is still a good representation of financial burden on future working generations.  Nevertheless, it should be aware that using OADR as a single measurement of financial burden might overlook some important economic factors. Some have suggested to use adjusted OADR that considers those important factors in analyzing pension finance [@HY12]. This paper emphasizes how to undertake a better forecasting of economic burden for a given policy rather than the better choice of measurements. Therefore, the simple version of OADR was chosen to illustrate my approach (i.e., without adjustment). There is a ground to extend the approach to incorporate more complex measurement which takes into account of other important economic factors.

@@

```{r changespension, echo=FALSE, fig.cap="Government's adjustments on pension age scheme.", fig.height=4, fig.width=8, message=FALSE, warning=FALSE, messages=FALSE}
year <- 2065-2007
proposed <- ts(pmin(70,
         c(rep(65,2017-2008), 65 + cumsum(rep(c(0.5,0),55))))[1:50],
         start=2008)
approved <- pmin(proposed,67)
current <- pmin(proposed,65)
cols <- c(current='#1b9e77',approved='#d95f02',proposed='#7570b3')
ggplot() +
  autolayer(proposed, series='proposed') +
  autolayer(approved, series='approved') +
  autolayer(current, series='current') +
  xlab("Year") + ylab("Age of retirement") +
  scale_color_manual(values=cols, name=" ",
                     breaks=c("current","approved","proposed"),
                     labels=c("Pension age at 65","Approved change","Proposed further change"))
```

@@

```{r getdata}
############################
## GET ALL THE DATA
############################

@@

```{r smoothing}
################################################
################ SMOOTHING STEP ################
################################################

@@

```{r models}
################################################
################ MODELLING STEP ################
################################################

@@

```{r componentsfertility, fig.cap="Fitted basis functions and coefficients for Australian fertility rates."}
par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
plot(fertf, 'c', comp=2)
```

@@

```{r simulation}
############### SIMULATION STEP #################
# This takes a while!
#aus.sim <- pop.sim(mortf, fertf, migf, firstyearpop=pop2010, N=2000)
#save(aus.sim,file='aussim.rda')
load("aussim.rda")

@@

```{r totalpopulationbyyears, fig.cap="Fifth-year forecasts of total population for each sex, along with 80% prediction intervals. The actually population from 1950 to 2010 is shown using circles.", fig.height=4}
par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
par(mfrow=c(1,2))
fy <- 2012+(1:50)
plot(1921:2012, colSums(pop2010$pop$female)/1e6,
     xlim=c(1975,2030),
     ylim=range(c(7,17),na.rm=TRUE),
     xlab="Year",
     ylab="Population (millions)",
     main="Total females")
polygon(c(fy,rev(fy)),
        c(totfpop.q[2,],rev(totfpop.q[4,])),
        col="yellow",border=FALSE)
lines(fy,totfpop.q[3,],lwd=2,col="blue")

@@

```{r totalpopulationin2028, fig.cap="Forecast population pyramid for 2028, along with 80% prediction intervals. The actual pyramid for 2010 is shown using dashed lines."}
par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(4.3,3.6,3.0,3.3))
plot(c(0,100),c(0,100),type="n", main="Forecast population: 2028",yaxt="n",
    xlab="Population ('000)",ylab="Age",xaxt="n",xlim=c(-1,1)*max(abs(popm.hi[,20]),abs(popf.hi[,20]))/1e3)
mtext("Male                       Female")
abline(v=0)
axis(1,at=seq(-300,0,by=50),labels=rev(seq(0,300,by=50)))
axis(1,at=seq(50,300,by=50))
axis(2,at=seq(0,100,by=10),labels=seq(0,100,by=10))
axis(4,at=seq(0,100,by=10),labels=seq(0,100,by=10))
axis(2,at=seq(0,100,by=20),labels=seq(0,100,by=20))
axis(4,at=seq(0,100,by=20),labels=seq(0,100,by=20))
axis(4,at=50,line=1.5,labels="Age",col=2,tick=FALSE)
polygon(c(-popm.lo[,20],rev(-popm.hi[,20]))/1000,c(0:100,100:0),col="yellow",border=FALSE)
polygon(c(popf.lo[,20],rev(popf.hi[,20]))/1000,c(0:100,100:0),col="yellow",border=FALSE)
lines(-pop2010$pop$male[,ncol(pop2010$pop$male)]/1000,0:100,col=1,lty=2)
lines(pop2010$pop$female[,ncol(pop2010$pop$female)]/1000,0:100,col=1,lty=2)
text(100,60,"2010")
text(-100,60,"2010")
lines(-popm.mean[,20]/1000,0:100,col="blue",lwd=2)
lines(popf.mean[,20]/1000,0:100,col="blue",lwd=2)
```

@@

```{r OADRat65, fig.cap="Fifty-year forecasts of the old aged dependency ratio with assumption of pension age equal to 65 for both sexes, along with 80% prediction interval."}
oadp.f <- oadr(pop2010,aus.sim,current)
autoplot(oadp.f) + xlab("year") + ylab("ratio") + 
    ggtitle("Old age dependency ratio forecasts with pension age=65")
```

```{r OADRcurrent, fig.cap="Fifty-year forecasts of the old aged dependency ratio associated with a pension age equal to 65 (black line) and equal to the pension age scheme with approved change (red line) respectively, along with the desired OADR level in grey (i.e., mean forecast of OADR in 2014). The actual OADRs for 1950-2010 are shown in black."}
oadp.f.67 <- oadr(pop2010, aus.sim, pension.age=approved)
oadp.f <- oadr(pop2010,aus.sim,pension.age=current)
oadp.f.prop <- oadr(pop2010,aus.sim,pension.age=proposed)
autoplot(oadp.f$x) +
  geom_hline(yintercept=0.21, col="gray") +
  autolayer(oadp.f, series="current", alpha=0.5) +
  autolayer(oadp.f.67, series="approved", alpha=0.5) +
  autolayer(oadp.f.prop, series="proposed", alpha=0.5) +
  ylab("ratio") + xlab("Year") +
  ggtitle("Old-age dependency ratio forecasts") +
  scale_color_manual(values=cols, name=" ",
                     breaks=c("current","approved","proposed"),
                     labels=c("Pension age at 65","Approved change","Proposed further change"))
```

@@

```{r pension}
###find the optimal scheme of pension.age based on mean forecast
target <- oadp.f$mean[4]

@@

```{r pensionupper}
###find the upper boundary of pension ages whose prediction interval contains target    
pension.age.upper <- proposed
for(i in 6:50)
{
  current.upper <- oadr(pop2010,aus.sim,
                    pension.age=pension.age.upper)$upper[i]
  while ((current.upper > target) &
         (max(diff(pension.age.upper[(i-1):50]))<=(1-1/12)))   {
    pension.age.upper[i:50] <- pension.age.upper[i:50] + 1/12
    current.upper <- oadr(pop2010,aus.sim,
                    pension.age=pension.age.upper)$upper[i]
  }
}
oadp.f.upper<- oadr(pop2010,
                    aus.sim,pension.age=pension.age.upper)
```

```{r pensionlower}
#####find the lower boundary of pension ages whose prediction interval contains target
pension.age.lower <- proposed
for(i in 6:20)
{
  current.lower <- oadr(pop2010,aus.sim,
                    pension.age=pension.age.lower)$lower[i]
  while ((current.lower > target) &
         (max(diff(pension.age.lower[(i-1):20]))<=(1-1/12)))   {
    pension.age.lower[i:20] <- pension.age.lower[i:20] + 1/12
    current.lower <- oadr(pop2010,aus.sim,
                  pension.age=pension.age.lower)$lower[i]
  }
}
oadp.f.lower<- oadr(pop2010, aus.sim,pension.age=pension.age.lower)
```

```{r pension forecast}
pension.fc <- structure(list(
  mean=ts(pension.age, start=2011),
  upper=pension.age.upper,
  lower=pension.age.lower,
  level=80,
  x=ts(rep(65,50),end=2010)),class='forecast')
```

@@

```{r pensionagetables}
year <- time(oadp.f$mean)
yr <- trunc(pension.age)
mth <- trunc((pension.age - trunc(pension.age)) * 12)
agechar <- paste(yr,"Y",ifelse(mth>0,paste("+",mth,"M",sep=""),""),sep="")

@@

```{r pensionagetables2}
year <- time(oadp.f$mean)
yr <- trunc(pension.age.lower)
mth <- trunc((pension.age.lower - trunc(pension.age.lower)) * 12)
age.lo <- paste(yr,"Y",ifelse(mth>0,paste("+",mth,"M",sep=""),""),sep="")
yr <- trunc(pension.age.upper)
mth <- trunc((pension.age.upper - trunc(pension.age.upper)) * 12)
age.up <- paste(yr,"Y",ifelse(mth>0,paste("+",mth,"M",sep=""),""),sep="")

@@

```{r OADRwithoptimal, fig.cap="The mean forecast of old aged dependency ratio associated with target pension age scheme (listed in Table 1) shown in blue, along with the desired OADR level in red (i.e., mean forecast of OADR in 2014). The 80% prediction interval of the forecast of OADR is shown as blue shading."}
##plots of OADR with plausible pension age schemes 
autoplot(oadp.f.optimal$x) + 
  geom_hline(yintercept=oadp.f$mean[4], col='red') +
  autolayer(oadp.f.optimal) +
  ggtitle("OADR forecast estimates for optimal pension ages") + 
  xlab("year") + ylab("ratio") 
```

```{r OADRboundaries, fig.cap="The  80% prediction interval forecast of OADR (in blue) associated with upper and lower boundary of plausible pension age schemes (listed in Table.2), along with desired OADR level (in red)."}
p1 <- autoplot(oadp.f.upper$x) + 
  geom_hline(yintercept=oadp.f$mean[4], col='red') +
  #autolayer(oadp.f.upper) +
  ggtitle("OADR for upper boundary pension ages") + 
  xlab("Year") + ylab("ratio")  +
  ylim(0.07,0.27)
p2 <- autoplot(oadp.f.lower$x) + 
  geom_hline(yintercept=oadp.f$mean[4], col='red') +
  #autolayer(oadp.f.lower) +
  ggtitle("OADR for lower boundary pension ages") + 
  xlab("Year") + ylab("ratio") +
  ylim(0.07,0.27)
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r comparisonofpensionages, fig.cap="The comparison between the target pension age scheme (in green) and real pension age scheme that has been set or proposed by government, along with the 80% confidence interval of the target pension age shown in green shade. The red solid line indicates the changes to the pension age from 65 to 67, announced in 2009."}
ggplot() +
  autolayer(pension.fc, series="Pension age to give OADR=0.21") +
#  geom_hline(yintercept=65, color='black') +
  autolayer(ts(c(rep(65,7),c(65.5,65.5,66.5,66.5,67,67),rep(67,37))[1:50],
  start=2013), series="Approved change") +
  autolayer(ts(c(rep(65,50)),start=1990), series="Pension age at 65") +
  xlab("Year") + ylab("Age of retirement")
```

```{r ComparisonwithOADR22, fig.cap="The comparison between the target pension age scheme (in green) associated with desired level equal to current level plus $1%$ (i.e., 22%) and real pension age scheme that has been set or proposed by government, along with the 80% confidence interval of the target pension age shown in green shade."}
```

```{r OADR_realpensionage, fig.cap="The mean forecast of old aged dependency ratio associated with different pension age schemes under investigation, along with the desired OADR level in grey (i.e., mean forecast of OADR in 2014). The 80% prediction interval of the forecast of OADR is shown as shading."}
```

@@

```{r rollingtest, fig.cap="The mean absolute error of $h$-step-ahead forecast for $h=1,2,\\dots,25$."}
########Run a rolling forecast origin 
t <- 2010-1950+1 
H <- 25 ##forecast horizon for test
k <- 25  ##years of data required to make a forecast
oadrtest <- matrix(0,25,t-k-H)
for (i in 1:(t-k-H))
{
  ausmort.short <- extract.years(ausmort.sm, years=1949+i+(0:(k-1))) 
  ausfert.short <- extract.years(ausfert.sm, years=1949+i+(0:(k-1)))
  ausmig.short <- extract.years(ausmig.sm, years=1949+i+(0:(k-1)))
  pop.short <- extract.years(pop2010, years=1949+i+(0:k))
  mort.fit.s <- coherentfdm(ausmort.short)
  mortfs <- forecast(mort.fit.s , H)
  fert.fit.s <- fdm(ausfert.short)
  # Fix problem
#  miss <- is.na(fert.fit.s$residuals$y)
#  fert.fit.s$residuals$y[miss] <- 0
  # End fix
  fertfs <- forecast( fert.fit.s, H)
  mig.fit.s <- coherentfdm(ausmig.short)
  migfs <- forecast(mig.fit.s, H)
  aus.sim.s <- pop.sim(mortfs, fertfs, migfs, firstyearpop=pop.short, N=3)
  oadp.fs <- oadr(pop.short,aus.sim.s)
  oadrtest[,i] <- oadp.fs$mean
}
oadrtest
oadr.real<- oadr(pop2010,aus.sim)$x
oadr.real.mat <- oadr.error <- matrix(NA, 25, 12)
for(i in 1:25)
  oadr.real.mat[i,] <- oadr.real[(i+59):(59+21+i)]

oadr.error <- abs(oadrtest-oadr.real.mat)
oadrave <- apply(oadr.error,1,mean)
oadrave

>>>>

---
title: "Forecasting the old age dependency ratio to determine the best pension age"
author: "Rob J Hyndman, Yijun Zeng and Han Lin Shang"
date: "`r format(Sys.time(), '%-d %B %Y')`"
fontsize: 11pt
papersize: a4
fontfamily: kpfonts
bibliography: pensionage.bib
biblio-style: authoryear-comp
subparagraph: true
toc: false
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(demography)
library(addb)
library(ggplot2)
source("functions-v4.R")
```

@@

In this paper, we use the old aged dependency ratio (OADR\footnote{All abbreviations have been summarized in Appendix A.}) as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working age population to finance pensions and health costs [@ARW07; @IGR02; @IGR07].  The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age [@ARW07]. This can be expressed using the formula below:
\begin{equation}
  OADR=\frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those projections to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method [@HU07] and the product-ratio method [@HBY13] to forecast a population age structure.  Forecasts of population structure have been built on forecasts of age and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age and gender-specific population structure.

Using these predicted population paths, we can calculate the average of OADR and predict the levels of old age dependency that correspond to different pension age schemes. By testing different pension age schemes, we aim to find the pension age with the desired OADR outcome. Therefore, other contributions of this research are (1) to examine the effect of adjusting pension ages by comparing OADR before and after adjustment; and (2) to suggest a suitable pension age scheme to ensure that the economic burden will remain relatively constant at a desired level in the future. It should be noted that the desired level has been set at 21%, which is the estimated current level of OADR in 2014. This level was chosen because, while the current financial burden on tax payers is affordable, it is approaching the limits of affordability. However, whether or not the current level is the most appropriate level to Australia remains to be examined and it is not the focus of this paper. This paper concentrates on providing a framework for policy makers to find a target pension age for a given desired level of OADR.

Some may argue that using OADR as a measure of financial burden for the next generation could be biased, with a large number of retirees able to rely on self-financed retirement income, such as superannuation. However, 80% of Australians over pension age fully or partly rely on the age pension [@Power14]. Therefore, the proportion of people over working age is still a good representation of financial burden on future working generations.  Nevertheless, using OADR as a single measurement of financial burden may overlook some important economic factors. @HY12 have suggested the use of an adjusted OADR that takes account of some of these factors in analyzing pension finance [@HY12]. This paper emphasizes how to undertake a better forecasting of economic burden for a given policy rather than the better choice of measurements. Therefore, the simple version of OADR was chosen to illustrate my approach (i.e., without adjustment). There is a ground to extend the approach to incorporate more complex measurement which takes into account of other important economic factors.

@@

```{r changespension, echo=FALSE, fig.cap="Government's adjustments on pension age scheme.", fig.height=4, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
year <- 2065-2007
proposed <- ts(pmin(70,
         c(rep(65,2017-2008), 65 + cumsum(rep(c(0.5,0),55))))[1:50],
         start=2008)
approved <- pmin(proposed,67)
current <- pmin(proposed,65)
cols <- c(current='#1b9e77',approved='#d95f02',proposed='#7570b3')
ggplot() +
  autolayer(proposed, series='proposed') +
  autolayer(approved, series='approved') +
  autolayer(current, series='current') +
  xlab("Year") + ylab("Age of retirement") +
  scale_color_manual(values=cols, name=" ",
                     breaks=c("current","approved","proposed"),
                     labels=c("Pension age at 65","Approved change","Proposed further change"))
```

@@

```{r getdata, cache=TRUE}
############################
## GET ALL THE DATA
############################

@@

```{r smoothing, cache=TRUE}
################################################
################ SMOOTHING STEP ################
################################################

@@

```{r models, cache=TRUE}
################################################
################ MODELLING STEP ################
################################################

@@

```{r componentsfertility, fig.cap="Fitted basis functions and coefficients for Australian fertility rates.", cache=TRUE}
par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
plot(fertf, 'c', comp=2)
```

@@

```{r simulation, cache=TRUE}
############### SIMULATION STEP #################
# This takes a while!
#aus.sim <- pop.sim(mortf, fertf, migf, firstyearpop=pop2010, N=2000)
#save(aus.sim,file='aussim.rda')
load("aussim.rda")

@@

```{r totalpopulationbyyears, fig.cap="Fifth-year forecasts of total population for each sex, along with 80% prediction intervals. The actually population from 1950 to 2010 is shown using circles.", fig.height=4, cache=TRUE}
par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
par(mfrow=c(1,2))
fy <- 2012+(1:50)
plot(1921:2012, colSums(pop2010$pop$female)/1e6,
     xlim=c(1975,2030),
     ylim=range(c(7,17),na.rm=TRUE),
     xlab="Year",
     ylab="Population (millions)",
     main="Total females")
polygon(c(fy,rev(fy)),
        c(totfpop.q[2,],rev(totfpop.q[4,])),
        col="yellow",border=FALSE)
lines(fy,totfpop.q[3,],lwd=2,col="blue")

@@

```{r totalpopulationin2028, fig.cap="Forecast population pyramid for 2028, along with 80% prediction intervals. The actual pyramid for 2010 is shown using dashed lines.", cache=TRUE}
par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(4.3,3.6,3.0,3.3))
plot(c(0,100),c(0,100),type="n", main="Forecast population: 2028",yaxt="n",
    xlab="Population ('000)",ylab="Age",xaxt="n",xlim=c(-1,1)*max(abs(popm.hi[,20]),abs(popf.hi[,20]))/1e3)
mtext("Male                       Female")
abline(v=0)
axis(1,at=seq(-300,0,by=50),labels=rev(seq(0,300,by=50)))
axis(1,at=seq(50,300,by=50))
axis(2,at=seq(0,100,by=10),labels=seq(0,100,by=10))
axis(4,at=seq(0,100,by=10),labels=seq(0,100,by=10))
axis(2,at=seq(0,100,by=20),labels=seq(0,100,by=20))
axis(4,at=seq(0,100,by=20),labels=seq(0,100,by=20))
axis(4,at=50,line=1.5,labels="Age",col=2,tick=FALSE)
polygon(c(-popm.lo[,20],rev(-popm.hi[,20]))/1000,c(0:100,100:0),col="yellow",border=FALSE)
polygon(c(popf.lo[,20],rev(popf.hi[,20]))/1000,c(0:100,100:0),col="yellow",border=FALSE)
lines(-pop2010$pop$male[,ncol(pop2010$pop$male)]/1000,0:100,col=1,lty=2)
lines(pop2010$pop$female[,ncol(pop2010$pop$female)]/1000,0:100,col=1,lty=2)
text(100,60,"2010")
text(-100,60,"2010")
lines(-popm.mean[,20]/1000,0:100,col="blue",lwd=2)
lines(popf.mean[,20]/1000,0:100,col="blue",lwd=2)
```

@@

```{r OADRat65, fig.cap="Fifty-year forecasts of the old aged dependency ratio with assumption of pension age equal to 65 for both sexes, along with 80% prediction interval.", cache=TRUE}
oadp.f <- oadr(pop2010,aus.sim,current)
autoplot(oadp.f) + xlab("year") + ylab("ratio") + 
    ggtitle("Old age dependency ratio forecasts with pension age=65")
```

```{r OADRcurrent, fig.cap="Fifty-year forecasts of the old aged dependency ratio associated with a pension age equal to 65 (black line) and equal to the pension age scheme with approved change (red line) respectively, along with the desired OADR level in grey (i.e., mean forecast of OADR in 2014). The actual OADRs for 1950-2010 are shown in black.", cache=TRUE}
oadp.f.67 <- oadr(pop2010, aus.sim, pension.age=approved)
oadp.f <- oadr(pop2010,aus.sim,pension.age=current)
oadp.f.prop <- oadr(pop2010,aus.sim,pension.age=proposed)
autoplot(oadp.f$x) +
  geom_hline(yintercept=0.21, col="gray") +
  autolayer(oadp.f, series="current", alpha=0.5) +
  autolayer(oadp.f.67, series="approved", alpha=0.5) +
  autolayer(oadp.f.prop, series="proposed", alpha=0.5) +
  ylab("ratio") + xlab("Year") +
  ggtitle("Old-age dependency ratio forecasts") +
  scale_color_manual(values=cols, name=" ",
                     breaks=c("current","approved","proposed"),
                     labels=c("Pension age at 65","Approved change","Proposed further change"))
```

@@

```{r pension, cache=TRUE}
###find the optimal scheme of pension.age based on mean forecast
target <- oadp.f$mean[4]

@@

```{r pensionupper, cache=TRUE}
###find the upper boundary of pension ages whose prediction interval contains target    
pension.age.upper <- proposed
for(i in 6:50)
{
  current.upper <- oadr(pop2010,aus.sim,
                    pension.age=pension.age.upper)$upper[i]
  while ((current.upper > target) &
         (max(diff(pension.age.upper[(i-1):50]))<=(1-1/12)))   {
    pension.age.upper[i:50] <- pension.age.upper[i:50] + 1/12
    current.upper <- oadr(pop2010,aus.sim,
                    pension.age=pension.age.upper)$upper[i]
  }
}
oadp.f.upper<- oadr(pop2010,
                    aus.sim,pension.age=pension.age.upper)
```

```{r pensionlower, cache=TRUE}
#####find the lower boundary of pension ages whose prediction interval contains target
pension.age.lower <- proposed
for(i in 6:20)
{
  current.lower <- oadr(pop2010,aus.sim,
                    pension.age=pension.age.lower)$lower[i]
  while ((current.lower > target) &
         (max(diff(pension.age.lower[(i-1):20]))<=(1-1/12)))   {
    pension.age.lower[i:20] <- pension.age.lower[i:20] + 1/12
    current.lower <- oadr(pop2010,aus.sim,
                  pension.age=pension.age.lower)$lower[i]
  }
}
oadp.f.lower<- oadr(pop2010, aus.sim,pension.age=pension.age.lower)
```

```{r pension forecast, cache=TRUE}
pension.fc <- structure(list(
  mean=ts(pension.age, start=2011),
  upper=pension.age.upper,
  lower=pension.age.lower,
  level=80,
  x=ts(rep(65,50),end=2010)),class='forecast')
```

@@

```{r pensionagetables, cache=TRUE}
year <- time(oadp.f$mean)
yr <- trunc(pension.age)
mth <- trunc((pension.age - trunc(pension.age)) * 12)
agechar <- paste(yr,"Y",ifelse(mth>0,paste("+",mth,"M",sep=""),""),sep="")

@@

```{r pensionagetables2, cache=TRUE}
year <- time(oadp.f$mean)
yr <- trunc(pension.age.lower)
mth <- trunc((pension.age.lower - trunc(pension.age.lower)) * 12)
age.lo <- paste(yr,"Y",ifelse(mth>0,paste("+",mth,"M",sep=""),""),sep="")
yr <- trunc(pension.age.upper)
mth <- trunc((pension.age.upper - trunc(pension.age.upper)) * 12)
age.up <- paste(yr,"Y",ifelse(mth>0,paste("+",mth,"M",sep=""),""),sep="")

@@

```{r OADRwithoptimal, fig.cap="The mean forecast of old aged dependency ratio associated with target pension age scheme (listed in Table 1) shown in blue, along with the desired OADR level in red (i.e., mean forecast of OADR in 2014). The 80% prediction interval of the forecast of OADR is shown as blue shading.", cache=TRUE}
##plots of OADR with plausible pension age schemes 
autoplot(oadp.f.optimal$x) + 
  geom_hline(yintercept=oadp.f$mean[4], col='red') +
  autolayer(oadp.f.optimal) +
  ggtitle("OADR forecast estimates for optimal pension ages") + 
  xlab("year") + ylab("ratio") 
```

```{r OADRboundaries, fig.cap="The  80% prediction interval forecast of OADR (in blue) associated with upper and lower boundary of plausible pension age schemes (listed in Table.2), along with desired OADR level (in red).", cache=TRUE}
p1 <- autoplot(oadp.f.upper$x) + 
  geom_hline(yintercept=oadp.f$mean[4], col='red') +
  #autolayer(oadp.f.upper) +
  ggtitle("OADR for upper boundary pension ages") + 
  xlab("Year") + ylab("ratio")  +
  ylim(0.07,0.27)
p2 <- autoplot(oadp.f.lower$x) + 
  geom_hline(yintercept=oadp.f$mean[4], col='red') +
  #autolayer(oadp.f.lower) +
  ggtitle("OADR for lower boundary pension ages") + 
  xlab("Year") + ylab("ratio") +
  ylim(0.07,0.27)
gridExtra::grid.arrange(p1,p2, ncol=2)
```

```{r comparisonofpensionages, fig.cap="The comparison between the target pension age scheme (in green) and real pension age scheme that has been set or proposed by government, along with the 80% confidence interval of the target pension age shown in green shade. The red solid line indicates the changes to the pension age from 65 to 67, announced in 2009.", cache=TRUE}
ggplot() +
  autolayer(pension.fc, series="Pension age to give OADR=0.21") +
#  geom_hline(yintercept=65, color='black') +
  autolayer(ts(c(rep(65,7),c(65.5,65.5,66.5,66.5,67,67),rep(67,37))[1:50],
  start=2013), series="Approved change") +
  autolayer(ts(c(rep(65,50)),start=1990), series="Pension age at 65") +
  xlab("Year") + ylab("Age of retirement")
```

```{r ComparisonwithOADR22, fig.cap="The comparison between the target pension age scheme (in green) associated with desired level equal to current level plus $1%$ (i.e., 22%) and real pension age scheme that has been set or proposed by government, along with the 80% confidence interval of the target pension age shown in green shade.", cache=TRUE}
```

```{r OADR_realpensionage, fig.cap="The mean forecast of old aged dependency ratio associated with different pension age schemes under investigation, along with the desired OADR level in grey (i.e., mean forecast of OADR in 2014). The 80% prediction interval of the forecast of OADR is shown as shading.", cache=TRUE}
```

@@

```{r rollingtest, fig.cap="The mean absolute error of $h$-step-ahead forecast for $h=1,2,\\dots,25$.", warning=FALSE, message=FALSE, cache=TRUE}
########Run a rolling forecast origin 
t <- 2010-1950+1 
H <- 25 ##forecast horizon for test
k <- 25  ##years of data required to make a forecast
oadrtest <- matrix(0,25,t-k-H)
for (i in 1:(t-k-H))
{
  ausmort.short <- extract.years(ausmort.sm, years=1949+i+(0:(k-1))) 
  ausfert.short <- extract.years(ausfert.sm, years=1949+i+(0:(k-1)))
  ausmig.short <- extract.years(ausmig.sm, years=1949+i+(0:(k-1)))
  pop.short <- extract.years(pop2010, years=1949+i+(0:k))
  mort.fit.s <- coherentfdm(ausmort.short)
  mortfs <- forecast(mort.fit.s , H)
  fert.fit.s <- fdm(ausfert.short)
  # Fix problem
#  miss <- is.na(fert.fit.s$residuals$y)
#  fert.fit.s$residuals$y[miss] <- 0
  # End fix
  fertfs <- forecast( fert.fit.s, H)
  mig.fit.s <- coherentfdm(ausmig.short)
  migfs <- forecast(mig.fit.s, H)
  aus.sim.s <- pop.sim(mortfs, fertfs, migfs, firstyearpop=pop.short, N=3)
  oadp.fs <- oadr(pop.short,aus.sim.s)
  oadrtest[,i] <- oadp.fs$mean
}
oadrtest
oadr.real<- oadr(pop2010,aus.sim)$x
oadr.real.mat <- oadr.error <- matrix(NA, 25, 12)
for(i in 1:25)
  oadr.real.mat[i,] <- oadr.real[(i+59):(59+11+i)]

oadr.error <- abs(oadrtest-oadr.real.mat[,1:11])
oadrave <- apply(oadr.error,1,mean)
oadrave
